---
title: "Homework 3"
author: Sarah McLarnan
date: 2019-10-07
output: github_document
---

### Problem 1
```{r setup, include=FALSE}
library(tidyverse)
library(viridis)
knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
theme_set(theme_bw() + theme(legend.position = "bottom"))
```

```{r problem_1_data_load}
library(p8105.datasets)
data("instacart")
```
The `instacart` data set has `r nrow(instacart)` observations of `r ncol(instacart)` metrics. Each row corresponds to a single item in an order. The columns contain information including the item ordered (`product_id`, `product_name`), the department (`department_id`, `department`) and aisle (`aisle_id`, `aisle`) it is from, the day of the week (`order_dow`) and time (`order_hour_of_day`) it was ordered, the order in which the item was added to the order (`add_to_cart_order`) in addition to other metrics. For example the first item in the table is Bulgarian Yogurt from the yogurt aisle in the dairy and eggs department. The item was ordered at 10 a.m. on a Wednesday, it was the first item the shopper added to their order and it had been 9 days since that shopper's previous order. 

```{r problem_1_aisles}
n_distinct(pull(instacart, aisle_id))

instacart %>%
  group_by(aisle) %>%
  summarize(n = n()) %>%
  mutate(aisle_rank = rank(desc(n))) %>%
  filter(aisle_rank <= 3)
```
There are 134 aisles. The top three aisles with the most products ordered from are fresh vegetables, fresh fruits, and packaged vegetables and fruits. 

```{r problem_1_aisle_graph}
instacart %>%
  group_by(aisle) %>%
  summarize(n = n()) %>%
  filter(n > 10000) %>%
  ggplot(aes(x = aisle, y = n)) + 
  geom_col() +
  theme(axis.text.x = element_text(angle = 90)) +
  labs(x = "Aisle", y = "Number of Products Ordered", title = "Number of Products Ordered from Aisle")
```
The graph of number of products ordered from each aisle shows the three aisles with the most products ordered were fresh fruits (~150,000) and fresh vegetables (~150,000) and packged fruits and vegetables (~77,000)

```{r problem_1_product_table}
instacart %>%
  group_by(aisle, product_name) %>%
  summarize(n = n()) %>%
  mutate(product_rank = rank(desc(n))) %>%
  filter(product_rank <= 3) %>%
  filter(aisle == "baking ingredients" | aisle == "dog food care" | aisle == "packaged vegetables fruits") %>%
  arrange(aisle, product_rank) %>%
  knitr::kable()
```
The top three products ordered in baking ingrendients were light brown sugar, pure baking soda, and cane sugar. In dog food care the top three products were snack sticks chicken and rice recipe dog treats, organix chicken and brown rice recipe and small dog biscuits. The top three products ordred in packaged vegetables and fruits were organic baby spinach, organic raspberries, and organic blueberries. 

```{r problem_1_product_time}
instacart %>%
  filter(product_name == "Coffee Ice Cream" | product_name == "Pink Lady Apples") %>%
  group_by (order_dow, product_name) %>%
  summarize (mean_time = mean(order_hour_of_day)) %>%
  pivot_wider (names_from = order_dow, values_from = mean_time) %>%
  knitr::kable()
```
The table of average time that each pink lady apples and coffe ice cream were ordered show that coffee ice cream is generally ordered later in the day than pink lady apples. Both products were ordered, on average, between late afternoon and mid afternoon.

### Problem 2
```{r problem_2_data_clean}
data("brfss_smart2010")

brfss_smart2010 = 
  brfss_smart2010 %>%
  janitor::clean_names() %>%
  filter(topic == "Overall Health") %>%
  filter(response == "Excellent" | response == "Very good" | response == "Good" | response == "Fair" | response == "Poor") %>%
  mutate(response = factor(response), 
         response = forcats::fct_relevel(response, c("Poor", "Fair", "Good", "Very good", "Excellent"))) 
```
```{r problem_2_multiple_location}
brfss_smart2010 %>%
  filter(year == "2002") %>%
  group_by(locationabbr) %>%
  summarize(unique_location = length(unique(locationdesc))) %>%
  filter(unique_location >= 7)

brfss_smart2010 %>%
  filter(year == "2010") %>%
  group_by(locationabbr) %>%
  summarize(unique_location = length(unique(locationdesc))) %>%
  filter(unique_location >= 7)
```
```{r problem_2_state_graph}
brfss_smart2010 %>%
  filter(response == "Excellent") %>%
  group_by(locationabbr, year) %>%
  summarize(data_value_mean = mean(data_value)) %>%
  ggplot(aes(x = year, y = data_value_mean, group = locationabbr, color = locationabbr)) +
  geom_line() +
  labs(
        x = "Year", 
        y = "Mean Data Value", 
        title = "Mean Data Value for Excellent Responses by Year for U.S. States"
      )
```
```{r problem_2_NY_graph}
brfss_smart2010 %>%
  filter(locationabbr == "NY", year == "2006" | year == "2010") %>%
  ggplot(aes(x = response, y = data_value)) +
        geom_col() +
        facet_grid(~year) +
        labs(x = "Response", y = "Data Value", title = "Data Value by Response at NY locations in 2006 and 2010")
```
```{r problem_3_data_tidy}
accel_data = read_csv("./data/accel_data.csv") 

accel_data = pivot_longer(accel_data, activity.1:activity.1440, names_to = "activity_minute", values_to = "activity_count") %>%
  separate(activity_minute, into = c("activity", "minute")) %>%
  select(-activity) %>%
  mutate(
    minute = as.numeric(minute),
    day = as.factor(day),
    day = forcats::fct_relevel(day, c("Sunday", "Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday")),
    day_type  = if_else(day == "Saturday" | day == "Sunday", "weekend", "weekday"))
```
```{r problem_3_activity_count}
accel_data %>%
  group_by(day_id) %>%
  summarize(total_activity = sum(activity_count)) %>%
  knitr::kable()
```
```{r problem_3_graph}
accel_data %>%
  ggplot(aes(x = minute, y = activity_count, color = day)) + 
  geom_line(alpha = .3) +
  scale_x_continuous(
    breaks = c(60, 120, 180, 240, 300, 360, 420, 480, 540, 600, 660, 720, 780, 840, 900, 960, 1020, 1080, 1140, 1200, 1260, 1320, 1380, 1440), 
    labels = c("1:00", "2:00", "3:00", "4:00", "5:00", "6:00", "7:00", "8:00", "9:00", "10:00", "11:00", "12:00", "13:00", "14:00", "15:00", "16:00", "17:00", "18:00", "19:00", "20:00", "21:00", "22:00", "23:00", "24:00")) +
  labs(x = "Hour of Day", y = "Activity Count", title = "24 Hour Activity Time Course by Day of Week") +
  theme(axis.text.x = element_text(angle = 90))
```




